Supplementary Content

### 2022-02-02 experiments/2_supplement/supplement.Rmd
working on the supplement

---
title: Supplementary Content
subtitle: Functional diversity ... Predictions of a model of an oxic-anoxic microbial ecosystem 
authors:
  - name: Owen L Petchey
    department: Department of Evolutionary Biology and Environmental Studies
    affiliation: University of Zurich
    location: Zurich 8057, Switzerland
    email: owen.petchey@ieu.uzh.ch
  - name: Rainer Krug
    department: Department of Evolutionary Biology and Environmental Studies
    affiliation: University of Zurich
    location: Zurich 8057, Switzerland
    email: rainer.krug@ieu.uzh.ch
  - name: Any Others
    department: Department of Evolutionary Biology and Environmental Studies
    affiliation: University of Zurich
    location: Zurich 8057, Switzerland
    email: any.others@ieu.uzh.ch
  - name: Marcel Suleiman
    department: Department of Evolutionary Biology and Environmental Studies
    affiliation: University of Zurich
    location: Zurich 8057, Switzerland
    email: marcel.suleiman@ieu.uzh.ch

bibliography: references.bib
biblio-style: unsrt
output: 
  rticles::arxiv_article:
    keep_tex: true
---

```{r include = FALSE, echo = FALSE}
knitr::opts_chunk$set(include = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      cache = FALSE)
```

```{r echo = FALSE}
rm(list = ls())
library(tidyverse)
library(microxanox)
library(patchwork)
library(here)
source(here("r/various_useful_functions.r"))
colfunc_CB <- colorRampPalette(c("#024F17", "#B5FFC9"))
colfunc_SB <- colorRampPalette(c("#7D1402", "#FCBEB3"))
colfunc_PB <- colorRampPalette(c("#6E0172", "#F9AEFC"))
zero <- 0
unity <- 1
```

@@

...

# Visualising among strain diversity

The method for creating among strain diversity is explained in the main report. Here we show three visualisations, one for nine strains, one for three, and one for two.

@@

## Nine strains

```{r nine-strains, echo = FALSE}

num_CB_strains <- 9
num_PB_strains <- 9
num_SB_strains <- 9
var_expt <- create_diversity_factorial(
  zero = zero, unity = unity,
  num_div_treatment_levels = num_div_treatment_levels,
  CB_gmax_div = CB_gmax_div, CB_h_div = CB_h_div,
  SB_gmax_div = SB_gmax_div, SB_h_div = SB_h_div,
  PB_gmax_div = PB_gmax_div, PB_h_div = PB_h_div,
  default_9strain = new_strain_parameter(
    n_CB = num_CB_strains,
    n_SB = num_SB_strains,
    n_PB = num_PB_strains,
    values_initial_state =  "bush_ssfig3"
  )
)
display_diversity(4,
  var_expt = var_expt,
  num_CB_strains = num_CB_strains,
  num_SB_strains = num_SB_strains,
  num_PB_strains = num_PB_strains
)
```

## Three strains

```{r three-strains, echo = FALSE}

num_CB_strains <- 3
num_PB_strains <- 3
num_SB_strains <- 3
var_expt <- create_diversity_factorial(
  zero = zero, unity = unity,
  num_div_treatment_levels = num_div_treatment_levels,
  CB_gmax_div = CB_gmax_div, CB_h_div = CB_h_div,
  SB_gmax_div = SB_gmax_div, SB_h_div = SB_h_div,
  PB_gmax_div = PB_gmax_div, PB_h_div = PB_h_div,
  default_9strain = new_strain_parameter(
    n_CB = num_CB_strains,
    n_SB = num_SB_strains,
    n_PB = num_PB_strains,
    values_initial_state =  "bush_ssfig3"
  )
)
display_diversity(4,
  var_expt = var_expt,
  num_CB_strains = num_CB_strains,
  num_SB_strains = num_SB_strains,
  num_PB_strains = num_PB_strains
)
```

## Two strains

```{r two-strains, echo = FALSE}

num_CB_strains <- 2
num_PB_strains <- 2
num_SB_strains <- 2
var_expt <- create_diversity_factorial(
  zero = zero, unity = unity,
  num_div_treatment_levels = num_div_treatment_levels,
  CB_gmax_div = CB_gmax_div, CB_h_div = CB_h_div,
  SB_gmax_div = SB_gmax_div, SB_h_div = SB_h_div,
  PB_gmax_div = PB_gmax_div, PB_h_div = PB_h_div,
  default_9strain = new_strain_parameter(
    n_CB = num_CB_strains,
    n_SB = num_SB_strains,
    n_PB = num_PB_strains,
    values_initial_state =  "bush_ssfig3"
  )
)
display_diversity(4,
  var_expt = var_expt,
  num_CB_strains = num_CB_strains,
  num_SB_strains = num_SB_strains,
  num_PB_strains = num_PB_strains
)
```

# Effect of number of strains

May not need this section, as number of strains effect is shown in figure of main ms.

```{r eval = FALSE, echo = FALSE}

## Figure 2b-f ----

ss_9s <- readRDS(here("experiments/0_ss_finding/data/9_strain_SS_data_1e3.RDS"))
sort(unique(ss_9s$CB_var_gmax_s))
sort(unique(ss_9s$SB_var_gmax_s))
ss_result <- ss_9s %>%
  filter(abs(CB_var_gmax_s - 0) < 0.0001,
         abs(SB_var_gmax_s - 0) < 0.0001,)

#temp <- ss_result$ssfind_result[[1]] %>%

#ss_result <- ss_9s$ssfind_result[[25]]

temp <- ss_result$ssfind_result[[1]] %>%
  mutate(a = 10^a_O) %>%
  # arrange(a) %>%
  # mutate(direction = rep(c("up", "down"), nrow(ss_result)/2)) %>%
  #mutate(direction = ifelse(initial_N_CB == 1, "up", "down")) %>%
  # select(-initial_N_CB, -a_O) %>%
  gather(species, quantity, 2:(ncol(.) - 2)) %>%
  mutate(
    var_type = ifelse(grepl("B_", species), "Organism", "Substrate"),
    functional_group = case_when(
      str_detect(species, "CB_5") ~ "CB",
      str_detect(species, "SB_5") ~ "SB",
      str_detect(species, "PB_5") ~ "PB"
    ),
    log10_quantity = log10(quantity + 1)
  )

line_width <- 2

p1 <- temp %>%
  dplyr::filter(functional_group == "CB")  %>%
  mutate(species = factor(species, levels = unique(species))) %>%
  ggplot(aes(x = log10(a), y = log10_quantity, col = species)) +
  geom_path(lwd = line_width) +
  ##ylab("log10(quantity [cells])") +
  ylab("Log(Quantity)") +
  xlab("Oxygen diffusivity") +
  scale_colour_manual(values = colfunc_CB(num_CB_strains)) +
  guides(colour = guide_legend(ncol = 3)) +
  theme(legend.position="none") +
  ggtitle("(b) Cyanobacteria")
p1# p1

patchwork <- p1 / p2 / p3 / p4 / p5
patchwork
#ggsave(here("experiments/1_figures_main_ms/figure_2b-f.pdf"))

## Figure 3 ----
## Stable state with diversity in all functional groups.

ss_9s <- readRDS(here("experiments/0_ss_finding/data/9_strain_SS_data_1e3.RDS"))
sort(unique(ss_9s$CB_var_gmax_s))
sort(unique(ss_9s$SB_var_gmax_s))
ss_result <- ss_9s %>%
  filter(abs(CB_var_gmax_s - 0.028254848) < 0.0001,
         abs(SB_var_gmax_s - 0.084764545) < 0.0001)

temp <- ss_result$ssfind_result[[1]] %>%
  mutate(a = 10^a_O) %>%
  # arrange(a) %>%
  # mutate(direction = rep(c("up", "down"), nrow(ss_result)/2)) %>%
  #mutate(direction = ifelse(initial_N_CB == 1, "up", "down")) %>%
  # select(-initial_N_CB, -a_O) %>%
  gather(species, quantity, 2:(ncol(.) - 2)) %>%
  mutate(
    var_type = ifelse(grepl("B_", species), "Organism", "Substrate"),
    functional_group = case_when(
      str_detect(species, "CB_") ~ "CB",
      str_detect(species, "SB_") ~ "SB",
      str_detect(species, "PB_") ~ "PB"
    ),
    log10_quantity = log10(quantity + 1)
  )

line_width <- 2

p1 <- temp %>%
  dplyr::filter(functional_group == "CB")  %>%
  mutate(species = factor(species, levels = unique(species))) %>%
  ggplot(aes(x = log10(a), y = log10_quantity, col = species)) +
  geom_path(lwd = line_width) +
  ##ylab("log10(quantity [cells])") +
  ylab("Log(Quantity)") +
  xlab("Oxygen diffusivity") +
  scale_colour_manual(values = colfunc_CB(num_CB_strains)) +
  guides(colour = guide_legend(ncol = 3)) +
  theme(legend.position="none") +
  ggtitle("(b) Cyanobacteria")
p1# p1

patchwork <- p1 / p2 / p3 / p4 / p5
patchwork
#ggsave(here("experiments/1_figures_main_ms/figure_3.pdf"))

## Figure 4 ----
## Shows effect of no diversity versus diversity in all functional groups.

ss_9s <- readRDS(here("experiments/0_ss_finding/data/9_strain_SS_data_1e3.RDS"))
sort(unique(ss_9s$CB_var_gmax_s))
sort(unique(ss_9s$SB_var_gmax_s))
ss_result_none <- ss_9s %>%
  filter(abs(CB_var_gmax_s - 0.0) < 0.0001,
         abs(SB_var_gmax_s - 0.0) < 0.0001)
ss_result_CB <- ss_9s %>%
  filter(abs(CB_var_gmax_s - 0.014958449) < 0.0001,
         abs(SB_var_gmax_s - 0.0) < 0.0001)
ss_result_SBPB <- ss_9s %>%
  filter(abs(CB_var_gmax_s - 0.0) < 0.0001,
         abs(SB_var_gmax_s - 0.044875347) < 0.0001)
ss_result_both <- ss_9s %>%
  filter(abs(CB_var_gmax_s - 0.014958449) < 0.0001,
         abs(SB_var_gmax_s - 0.044875347) < 0.0001)

temp_none <- ss_result_none$ssfind_result[[1]] %>%
  mutate(a = 10^a_O) %>%
  # arrange(a) %>%
  # mutate(direction = rep(c("up", "down"), nrow(ss_result)/2)) %>%
  #mutate(direction = ifelse(initial_N_CB == 1, "up", "down")) %>%
  # select(-initial_N_CB, -a_O) %>%
  gather(species, quantity, 2:(ncol(.) - 2)) %>%
  mutate(
    var_type = ifelse(grepl("B_", species), "Organism", "Substrate"),
    functional_group = case_when(
      str_detect(species, "CB_") ~ "CB",
      str_detect(species, "SB_") ~ "SB",
      str_detect(species, "PB_") ~ "PB"
    ),
    log10_quantity = log10(quantity + 1)
  )

temp_CB <- ss_result_CB$ssfind_result[[1]] %>%
  mutate(a = 10^a_O) %>%
  # arrange(a) %>%
  # mutate(direction = rep(c("up", "down"), nrow(ss_result)/2)) %>%
  #mutate(direction = ifelse(initial_N_CB == 1, "up", "down")) %>%
  # select(-initial_N_CB, -a_O) %>%
  gather(species, quantity, 2:(ncol(.) - 2)) %>%
  mutate(
    var_type = ifelse(grepl("B_", species), "Organism", "Substrate"),
    functional_group = case_when(
      str_detect(species, "CB_") ~ "CB",
      str_detect(species, "SB_") ~ "SB",
      str_detect(species, "PB_") ~ "PB"
    ),
    log10_quantity = log10(quantity + 1)
  )
temp_SBPB <- ss_result_SBPB$ssfind_result[[1]] %>%
  mutate(a = 10^a_O) %>%
  # arrange(a) %>%
  # mutate(direction = rep(c("up", "down"), nrow(ss_result)/2)) %>%
  #mutate(direction = ifelse(initial_N_CB == 1, "up", "down")) %>%
  # select(-initial_N_CB, -a_O) %>%
  gather(species, quantity, 2:(ncol(.) - 2)) %>%
  mutate(
    var_type = ifelse(grepl("B_", species), "Organism", "Substrate"),
    functional_group = case_when(
      str_detect(species, "CB_") ~ "CB",
      str_detect(species, "SB_") ~ "SB",
      str_detect(species, "PB_") ~ "PB"
    ),
    log10_quantity = log10(quantity + 1)
  )
temp_both <- ss_result_both$ssfind_result[[1]] %>%
  mutate(a = 10^a_O) %>%
  # arrange(a) %>%
  # mutate(direction = rep(c("up", "down"), nrow(ss_result)/2)) %>%
  #mutate(direction = ifelse(initial_N_CB == 1, "up", "down")) %>%
  # select(-initial_N_CB, -a_O) %>%
  gather(species, quantity, 2:(ncol(.) - 2)) %>%
  mutate(
    var_type = ifelse(grepl("B_", species), "Organism", "Substrate"),
    functional_group = case_when(
      str_detect(species, "CB_") ~ "CB",
      str_detect(species, "SB_") ~ "SB",
      str_detect(species, "PB_") ~ "PB"
    ),
    log10_quantity = log10(quantity + 1)
  )

line_width <- 2.5

## a
p1 <- temp_none %>%
  dplyr::filter(var_type == "Substrate", species == "O") %>%
  ggplot(aes(x = log10(a), y = log10_quantity, col = species)) +
  geom_path(lwd = line_width) +
  geom_path(data = filter(temp_CB, var_type == "Substrate", species == "O"),
            col = "black", lwd = line_width - 1.5) +
  ##ylab("log10(quantity [cells])") +
  ylab("Log(Quantity)") +
  xlab("Oxygen diffusivity") +
  theme(legend.position="none") +
  ggtitle("(a) No diversity versus diversity in only cyanobacteria") +
  geom_vline(xintercept = c(-8, 0), col = "grey", lwd = 3)
p2 <- temp_none %>%
  dplyr::filter(var_type == "Substrate", species == "O") %>%
  ggplot(aes(x = log10(a), y = log10_quantity, col = species)) +
  geom_path(lwd = line_width) +
  geom_path(data = filter(temp_SBPB, var_type == "Substrate", species == "O"),
            col = "black", lwd = line_width - 1.5) +
  ##ylab("log10(quantity [cells])") +
  ylab("Log(Quantity)") +
  xlab("Oxygen diffusivity") +
  theme(legend.position="none") +
  ggtitle("(b) No diversity versus diversity in only sulfur bacteria") +
  geom_vline(xintercept = c(-8, 0), col = "grey", lwd = 3)
p3 <- temp_none %>%
  dplyr::filter(var_type == "Substrate", species == "O") %>%
  ggplot(aes(x = log10(a), y = log10_quantity, col = species)) +
  geom_path(lwd = line_width) +
  geom_path(data = filter(temp_both, var_type == "Substrate", species == "O"),
            col = "black", lwd = line_width - 1.5) +
  ##ylab("log10(quantity [cells])") +
  ylab("Log(Quantity)") +
  xlab("Oxygen diffusivity") +
  theme(legend.position="none") +
  ggtitle("(c) No diversity versus diversity all functional groups") +
  geom_vline(xintercept = c(-8, 0), col = "grey", lwd = 3)

p1 / p2 / p3

#ggsave(here("experiments/1_figures_main_ms/figure_4.pdf"),
#       width = 6, height = 5)

## Figure 5 ----
## Region of bistability by diversity and number of strains.

#ss_2s <- readRDS(here("experiments/0_ss_finding/data/2_strain_SS_data.RDS"))
#stab_2s <- readRDS(here("experiments/0_ss_finding/data/2_strain_stab_data.RDS")) %>%
#  mutate(num_strains = 2)
ss_3s <- readRDS(here("experiments/0_ss_finding/data/3_strain_SS_data_1e3.RDS"))
stab_3s <- readRDS(here("experiments/0_ss_finding/data/3_strain_stab_data_1e3.RDS")) %>%
  mutate(num_strains = 3)
#ss_6s <- readRDS(here("experiments/0_ss_finding/data/5-level factorial/6_strain_SS_data.RDS"))
#stab_6s <- readRDS(here("experiments/0_ss_finding/data/5-level factorial/6_strain_stab_data.RDS")) %>%
#  mutate(num_strains = 6)
ss_9s <- readRDS(here("experiments/0_ss_finding/data/9_strain_SS_data_1e3.RDS"))
stab_9s <- readRDS(here("experiments/0_ss_finding/data/9_strain_stab_data_1e3.RDS")) %>%
  mutate(num_strains = 9)

all_stab <- #stab_2s %>%
  bind_rows(stab_3s) %>%
  #bind_rows(stab_6s) %>%
  bind_rows(stab_9s)

all_stab_results_small <- select(all_stab_results, -7:-12)

all_stab_results %>%
  #filter(var_treat == "CB") %>%
  filter(Species == "O") %>%
  ggplot(aes(x = var_gmax,
             col = as.factor(num_strains))) +
  geom_line(aes(y = hyst_min_log), lwd = 1.5) +
  geom_line(aes(y = hyst_max_log), lwd = 1.5) +
  facet_wrap( ~ var_treat, scales = "free_y", nrow = 3) +
  xlab("Amount of trait variation\n[see text for units]") +
  ylab("Oxygen diffusivity\n[log10 uM per hour]") +
  labs(fill = "Variation in\nonly these\nfunctional groups") +
  coord_flip() +
  guides(col = guide_legend(title="Number of strains")) +
  theme(
    legend.position="top"
    #strip.background = element_blank(),
    #strip.text.x = element_blank()
  ) +
  geom_hline(yintercept = c(-8, 0), col = "grey", lwd = 3)

temp <- all_stab_results %>%
  #filter(var_treat == "CB") %>%
  filter(Species == "O",
         var_treat == "SB-PB",
         num_strains == 9) %>%
  ggplot(aes(x = var_gmax,
             col = as.factor(num_strains))) +
  geom_line(aes(y = hyst_min_log)) +
  geom_line(aes(y = hyst_max_log)) +
  facet_wrap( ~ var_treat, scales = "free_y", nrow = 3) +
  xlab("Amount of trait variation\n[see text for units]") +
  ylab("Oxygen diffusivity\n[log10 uM per hour]") +
  labs(fill = "Variation in\nonly these\nfunctional groups") +
  coord_flip() +
  guides(col = guide_legend(title="Number of strains")) +
  theme(
    legend.position="top"
    #strip.background = element_blank(),
    #strip.text.x = element_blank()
  ) +
  geom_hline(yintercept = c(-8, 0), col = "grey", lwd = 3)

## 
## 
```

# Robustness of stable state finding

We now discuss two approaches for finding how the stable state of the system varies along the gradient of oxygen diffusivity when there exists the possibility of multiple stable states. In the main article we report results of the method also described there. Briefly here, we made a separate simulations for each of many values of oxygen diffusivity, ran the simulations for long enough to ensure transients had passed, and recorded the state. We ran multiple simulations per value of oxygen diffusivity, and started them with different initial conditions (either low or high total cyanobacterial abundance), so as to check for presence of alternate stable states. This method used by [Bush et al 2017](https://www.nature.com/articles/s41467-017-00912-x). We hereafter term this "the replication method", since it is effectively achieved by having many independent replicates of the ecosystem, each with a different oxygen diffusivity, and also a separate replicate for each of the different chosen initial conditions.

Here we report the results of a second method. We will show some quantitative differences in results compared to the other method, but also show the results concerning diversity effects are qualitatively the same as obtained with the first method.

The second method involves two simulations for a particular system configuration (parameter set). In one simulation the oxygen diffusivity is *increased* in a step-wise fashion. In the other it is *decreased* in a step-wise fashion. That is, oxygen diffusivity is held at a constant level for long enough for steady state to be reach, that state is recorded, and then a slightly higher (or lower) oxygen diffusivity value is set. Hence, at that time point, the system is effectively started with initial conditions that are the state of the system in the previous time step. We hereafter term this "the temporal method".

The temporal method results in a larger region of bistability than the replication method (Figure \@ref(fig:tempmeth1)). That is, a greater extent of environmental change is required to switch the system to the other state (i.e. the region of bistability is greater). This occurs both during increasing and decreasing oxygen diffusivities. The difference in results occurs because, as mentioned above, the initial conditions differ between the two approaches. For example, in the temporal method, the oxygen concentration in the anoxic state is lower than the initial oxygen concentration in the replication approach, and **Owen guesses that** the sulfide concentration is higher **check that last statement**.

## Replication method

```{r repl_meth1, echo=FALSE}
repl_meth <- readRDS(here("experiments/0_ss_finding/replication_method/9_strains/data/ss_data_1e6_x2x6_factorial.RDS"))
temp_meth <- readRDS(here("experiments/0_ss_finding/temporal_method/data/9_strain_SS_data.RDS"))
## To be completed.
plot_ss_result1(repl_meth,
                result_index = 1,
                filename_prefix = NULL,
                save_image_file = FALSE)
```

## Temporal method

```{r temp_meth1, echo=FALSE}
plot_ss_result3(temp_meth, 1)
```

## Both methods overlaid

```{r both_meth1, echo=FALSE}
repl_method_1 <- repl_meth$ss_res[[1]] %>%
     mutate(time = NA,
            direction = ifelse(initial_N_CB == 1, "up", "down")) %>%
  select(names(temp_meth$ssfind_result[[1]]))

@@

The replication method shows the same effects of functional diversity as those revealed by the replication method:

-   Introducing functional diversity in all three functional groups increases the region of bistability relative to when there is no diversity within all functional groups)..
-   Introducing functional diversity in only the cyanobacteria function group delays (with respect to the oxygen diffusivity it occurs at) the switch to the anoxic state as oxygen diffusivity is reduced, but has no effect on the oxygen diffusivity at which the anoxic to oxic transition occurs .
-   Introducing function diversity in only the sulfur bacteria delays (with respect to the oxygen diffusivity it occurs at) the switch from anoxic to oxic, and advances the switch from oxic to anoxic (Figure \@ref(fig:tempmeth4).

```{r echo = FALSE}
stab_data <- readRDS(here("experiments/0_ss_finding/replication_method/9_strains/data/stab_data_1e6_x2x6_factorial.RDS"))

@@

all_stab_results_small <- select(all_stab_results, -7:-12)

>>>>

---
title: Supplementary Content
subtitle: Functional diversity ... Predictions of a model of an oxic-anoxic microbial ecosystem 
author: "Owen L Petchey, Rainer Krug, Marcel Suleiman, ???others???"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::html_document2:
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
---

```{r include = FALSE, echo = FALSE}
knitr::opts_chunk$set(include = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      cache = FALSE)
```

```{r echo = FALSE}
rm(list = ls())
library(tidyverse)
library(microxanox)
library(patchwork)
library(here)
source(here("r/various_useful_functions.r"))
source(here("experiments/1_figures_main_ms/ms_figure_functions.r"))

#print(fig_div_vs_o2diff_1strain_7row)

colfunc_CB <- colorRampPalette(c("#024F17", "#B5FFC9"))
colfunc_SB <- colorRampPalette(c("#7D1402", "#FCBEB3"))
colfunc_PB <- colorRampPalette(c("#6E0172", "#F9AEFC"))
zero <- 0
unity <- 1
```

@@

This document contains the Supporting Content for the report *?????????????*.

# Creating among strain diversity

Here we provide additional description of the method used to create strain variation within functional groups. The following should be read only after the relevant sections of the main text.

Take for example the cyanobacteria functional group, a reference maximum growth rate parameter $G_{max,CB}$ and a reference tolerance parameter $h_{SR,CB}$. Recall that amount of among strain variation is controlled by the two meta-parameters $Div_{G_{max,CB}}$ and $Div_{h_{SR,CB}}$. To create a tradeoff, the two meta-parameters were always of different sign. The among strain variation was calculated such that the growth rate of strain $i = 1$ was a factor $1/({2Div_{G_{max,CB}}})$ of the reference growth rate, and the growth rate of the $i = n$ strain was a factor $2Div_{G_{max,CB}}$ of the reference growth rate. The growth rate of the $i = {2,...n-1}$ other strains was the reference growth rate multiplied by a factor that was equally distributed between the factor of the $i = 1$ and $i = n$ strain. This implementation of the variation and trade-off has the assumption of a linear trade-off of log-log transformed among-strain variation. Hence, for example, with the reference growth rate $G_{max,CB} = 0.05$, among strain variation of $Div_{G_{max,CB}} = 1$, and nine strains ($n = 9$), the growth rates of the nine strains are 0.025, 0.03, 0.035, 0.042, 0.05, 0.059, 0.071, 0.084, 0.1. Put another way, the growth rate parameter of the $i$-th of the $n$ strains in a functional group was calculated as $G_{max,i} = D_i * G_{max}$ where $D_i$ is the $i$-th element of the set $D = {2f(x)Div_{G_{max}} |x ∈ N, x ≤ n}$, where $f(x) = (x − (n + 1)/2)/((n − 1)/2)$.

This procedure ensures that the range of trait values is independent of the number of strains. When there were only two strains, the two strains took the minimum and maximum traits values, as defined by the $Div$ variable for that trait. When there were three strains, one had the minimum trait value, one the maximum, and one the mean. With nine strains, one had the minimum trait value, one the maximum, one the mean, and three were equally spaced (on log scale) between the minimum and the mean, and the remaining three between the maximum and the mean.

We show three examples of the created variation among strains within functional groups, one for nine strains (Figure \@ref(fig:nine-strain-div)), one for three (Figure \@ref(fig:three-strain-div)), and one for two (Figure \@ref(fig:two-strain-div)). In each, three graphs shows the diversity among strains in the maximum growth rate trait and the tolerance trait, plotted against each other and therefore also showing the tradeoff between the two traits.

Notice that in all three cases, the range of trait variation represented by the strains is the same, even though the number of strains differs (a consequence is that the lower row of panels in Figures \@ref(fig:nine-strain-div), \@ref(fig:three-strain-div), and \@ref(fig:two-strain-div) are identical). This approach allowed us to independently manipulate trait diversity and number of strains.

@@

All three examples have the following amounts of trait variation: $Div_{G_{max,CB}}$ = `r round(CB_gmax_div, 3)`, $Div_{h_{SR,CB}}$ = `r round(CB_h_div, 3)`, $Div_{G_{max,SB}}$ = `r round(SB_gmax_div, 3)`, $Div_{h_{O,SB}}$ = `r round(SB_h_div, 3)`, $Div_{G_{max,PB}}$ = `r round(PB_gmax_div, 3)`, $Div_{h_{O,PB}}$ = `r round(PB_h_div, 3)`. These are also the maximum amounts of trait variation explored in the simulations.

(ref:nine-strain-div) Trait variation with nine strains in each of the functional groups. In all panels colour indicates the functional group, darker shades show more tolerant-lower growth rate strains, lighter shades show less tolerant-higher growth rate strains.

```{r nine-strain-div, echo = FALSE,  include=TRUE, fig.cap='(ref:nine-strain-div)', fig.align="center", fig.height=3, fig.width=8}

num_CB_strains <- 9
num_PB_strains <- 9
num_SB_strains <- 9
var_expt <- create_diversity_factorial(
  zero = zero, unity = unity,
  num_div_treatment_levels = num_div_treatment_levels,
  CB_gmax_div = CB_gmax_div, CB_h_div = CB_h_div,
  SB_gmax_div = SB_gmax_div, SB_h_div = SB_h_div,
  PB_gmax_div = PB_gmax_div, PB_h_div = PB_h_div,
  default_9strain = new_strain_parameter(
    n_CB = num_CB_strains,
    n_SB = num_SB_strains,
    n_PB = num_PB_strains,
    values_initial_state =  "bush_ssfig3"
  )
)
display_diversity(4,
  var_expt = var_expt,
  num_CB_strains = num_CB_strains,
  num_SB_strains = num_SB_strains,
  num_PB_strains = num_PB_strains,
  which_graphs = 1
)
```

(ref:three-strain-div) Trait variation with three strains in each of the functional groups. Note that the three displayed strains are the 1st, 5th, and 9th strains from the nine strain example shown in Figure \@ref(fig:nine-strain-div). In all panels colour indicates the functional group, darker shades show more tolerant-lower growth rate strains, lighter shades show less tolerant-higher growth rate strains.

```{r three-strain-div, echo = FALSE,  include=TRUE, fig.cap='(ref:three-strain-div)', fig.height=3, fig.width=8, fig.align="center"}

num_CB_strains <- 3
num_PB_strains <- 3
num_SB_strains <- 3
var_expt <- create_diversity_factorial(
  zero = zero, unity = unity,
  num_div_treatment_levels = num_div_treatment_levels,
  CB_gmax_div = CB_gmax_div, CB_h_div = CB_h_div,
  SB_gmax_div = SB_gmax_div, SB_h_div = SB_h_div,
  PB_gmax_div = PB_gmax_div, PB_h_div = PB_h_div,
  default_9strain = new_strain_parameter(
    n_CB = num_CB_strains,
    n_SB = num_SB_strains,
    n_PB = num_PB_strains,
    values_initial_state =  "bush_ssfig3"
  )
)
display_diversity(4,
  var_expt = var_expt,
  num_CB_strains = num_CB_strains,
  num_SB_strains = num_SB_strains,
  num_PB_strains = num_PB_strains,
  which_graphs = 1
)
```

(ref:two-strain-div) Trait variation with two strains in each of the functional groups. Note that the two displayed strains are the 1st and 9th strains from the nine strain example shown in Figure \@ref(fig:nine-strain-div). In all panels colour indicates the functional group, darker shades show more tolerant-lower growth rate strains, lighter shades show less tolerant-higher growth rate strains.

```{r two-strain-div, echo = FALSE,  include=TRUE, fig.cap='(ref:two-strain-div)', fig.height=3, fig.width=8, fig.align="center"}

num_CB_strains <- 2
num_PB_strains <- 2
num_SB_strains <- 2
var_expt <- create_diversity_factorial(
  zero = zero, unity = unity,
  num_div_treatment_levels = num_div_treatment_levels,
  CB_gmax_div = CB_gmax_div, CB_h_div = CB_h_div,
  SB_gmax_div = SB_gmax_div, SB_h_div = SB_h_div,
  PB_gmax_div = PB_gmax_div, PB_h_div = PB_h_div,
  default_9strain = new_strain_parameter(
    n_CB = num_CB_strains,
    n_SB = num_SB_strains,
    n_PB = num_PB_strains,
    values_initial_state =  "bush_ssfig3"
  )
)
display_diversity(4,
  var_expt = var_expt,
  num_CB_strains = num_CB_strains,
  num_SB_strains = num_SB_strains,
  num_PB_strains = num_PB_strains,
  which_graphs = 1
)
```

We also show, in Figure \@ref(fig:growth-rates), the relationship between realised growth rate and the concentration of the respective inhibiting substance, e.g. sulfide for the cyanobacteria. In these graphs only the least and most tolerant strains are shown. The lines cross due to the tradeoff. At high concentrations of the inhibiting substance the more tolerant strain has higher realised growth rate, while at lower concentrations it is the less tolerant strain that has the higher realised growth rate.

(ref:growth-rates) Response of realised growth rate to variation in concentration of the inhibiting substrate for two strains, namely the most and the least tolerant strain. As always colour indicates the functional group, darker shades show more tolerant-lower growth rate strains, lighter shades show less tolerant-higher growth rate strains.

```{r growth-rates, echo = FALSE,  include=TRUE, fig.cap='(ref:growth-rates)', fig.height=3, fig.width=8, fig.align="center"}
display_diversity(4,
  var_expt = var_expt,
  num_CB_strains = num_CB_strains,
  num_SB_strains = num_SB_strains,
  num_PB_strains = num_PB_strains,
  which_graphs = 2
)
```

# Method for finding stable states

We now discuss two approaches for finding how the stable state of the system varies along the gradient of oxygen diffusivity when there exists the possibility of multiple stable states. In the main article we describe and report results of the method that we term the *Temporal method*.

Here we also describe and give results of what we term the *Replication method*. In this, we made a separate simulations for each of many values of oxygen diffusivity, ran the simulations for long enough to ensure transients had passed, and recorded the state. We ran multiple simulations per value of oxygen diffusivity, and started them with different initial conditions (either low or high total cyanobacterial abundance), so as to check for presence of alternate stable states. This method used by [Bush et al 2017](https://www.nature.com/articles/s41467-017-00912-x). We term it the replication method since it is achieved by having many independent replicates of the ecosystem, each with a different oxygen diffusivity, and also a separate replicate for each of the different chosen initial conditions.

As is shown in the figures below, the temporal method results in a larger region of bistability than the replication method, but only during increasing oxygen diffusivities. The difference in results occurs because, as mentioned above, the initial conditions differ between the two approaches. For example, in the temporal method, the oxygen concentration in the anoxic state is lower than the initial oxygen concentration in the replication approach, and **Owen guesses that** the sulfide concentration is higher **check that last statement**.

## Replication method, no diversity

In the following graph, the solid (down) line refers to the stable state when cyanobacteria begin at high abundance (favouring the oxic state). The dashed up line refers to when their initial abundnace is low (favouring the anoxic state).Stable states found with the replication method. Please note: i) that near vertical lines indicate the position of a tipping point and hence the system cannot exist on point of that part of the line; ii) the separatrix (the unstable equilibrium) is not shown.

```{r repl_meth1, echo=FALSE, eval = FALSE}
repl_meth <- readRDS(here("experiments/0_ss_finding/replication_method/9_strains/data/ss_data_1e6_x2x6_factorial.RDS"))
temp_meth <- readRDS(here("experiments/0_ss_finding/temporal_method/data/ss_data_9strains_waittime1e+06_event_definition_2.RDS"))
## To be completed.
plot_ss_result1(repl_meth,
                result_index = 1,
                filename_prefix = NULL,
                save_image_file = FALSE)
```

## Temporal method, no diversity

Same as for the previous graph.

```{r temp_meth1, echo=FALSE, eval = FALSE}
plot_ss_result3(temp_meth, 1)
```

## Both methods overlaid, no diversity

Stable states of the two methods overlaid, the temporal method with a dashed line, and the replication method with a solid line.

```{r both_meth1, echo=FALSE, eval = FALSE}
repl_method_1 <- repl_meth$ss_res[[1]] %>%
     mutate(time = NA,
            direction = ifelse(initial_N_CB == 1, "up", "down")) %>%
  select(names(temp_meth$ssfind_result[[1]]))

@@

The replication method shows the same effects of functional diversity as those revealed by the temporal method:

-   Introducing functional diversity in all three functional groups increases the region of bistability relative to when there is no diversity within all functional groups).
-   Introducing functional diversity in only the cyanobacteria function group delays (with respect to the oxygen diffusivity it occurs at) the switch to the anoxic state as oxygen diffusivity is reduced, but has no effect on the oxygen diffusivity at which the anoxic to oxic transition occurs .
-   Introducing function diversity in only the sulfur bacteria delays (with respect to the oxygen diffusivity it occurs at) the switch from anoxic to oxic, and advances the switch from oxic to anoxic.

These results are shown in the figure below, which shows results of the replication method.

```{r echo = FALSE, eval = FALSE}
stab_data <- readRDS(here("experiments/0_ss_finding/replication_method/9_strains/data/small_stab_data_1e6_x2x6_factorial.RDS"))

@@

```{r echo = FALSE, eval = FALSE}
all_stab_results %>%
  #filter(var_treat == "CB") %>%
  filter(Species == "SR") %>%
  ggplot(aes(x = var_gmax)) +
  #ggplot(aes(x = var_gmax,
  #           col = as.factor(num_strains))) +
  geom_line(aes(y = hyst_min_log), lwd = 2, alpha = 0.8) +
  geom_line(aes(y = hyst_max_log), lwd = 2, alpha = 0.8) +
  facet_wrap( ~ var_treat, scales = "free_y", nrow = 3) +
  xlab("Amount of trait variation\n[see text for units]") +
  ylab("Oxygen diffusivity\n[log10 uM per hour]") +
  labs(fill = "Variation in\nonly these\nfunctional groups") +
  coord_flip() +
  #guides(col = guide_legend(title="Number of strains")) +
  theme(
    legend.position="top"
    #strip.background = element_blank(),
    #strip.text.x = element_blank()
  ) +
  geom_hline(yintercept = c(-8, 0), col = "grey", lwd = 3)
